---
title: "PortfolioConstruction"
author: "Reng Chiz Der"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```
### Portfolio Construction Test
Please note that the portfolio used are the same across the two tasks (bootstrap and portfolio)
### Environment Setup 
```{r}
library(quadprog)
library(ROI)
library(ROI.plugin.glpk)
library(ROI.plugin.quadprog)
library(ROI.plugin.symphony)
library(quantmod)
library(zoo)
library(xts)
library(ggplot2)
library(PerformanceAnalytics)
library(PortfolioAnalytics)
```

### Getting data
```{r}
fetch_and_preprocess <- function(tickers) {
  stocks <- lapply(tickers, function(ticker) {
    stock_data <- getSymbols(ticker, src = 'yahoo', from = "2020-01-01", to = "2024-01-01", auto.assign = FALSE)
    colnames(stock_data) <- gsub(".*\\.", "", colnames(stock_data)) # Remove the prefix
    adjusted_data <- stock_data[, "Adjusted", drop = FALSE] # Select only the adjusted closing value
    if ("Adjusted" %in% colnames(stock_data)) {
      colnames(adjusted_data) <- ticker
      return(adjusted_data)
    } else {
      stop("Adjusted closing value not found for", ticker)
    }
  })
  
  # Preprocess the data
  for (i in seq_along(stocks)) {
    if (!is.null(stocks[[i]])) {
      # Calculate the number of missing values
      missing_values <- colSums(is.na(stocks[[i]]))
      
      # Replace missing values with the previous day's values
      stocks[[i]] <- na.locf(stocks[[i]])
    }
  }
  
  # Combine the data into a single xts object
  combined_data <- do.call(merge, stocks)
  return(combined_data)
}

# Example usage:
stocks <- c('AAPL', 'NVDA', 'AMD')
stock_data <- fetch_and_preprocess(stocks)

# S&P500, Emerging Market ETF
etfs <- c('SPY', 'SPEM')
etf_data <- fetch_and_preprocess(etfs)

# fixed income (bonds): IEF (treasury bond ETF), AGG (US Aggregate Bond ETF)
fis <- c('IEF', 'AGG')
fi_data <- fetch_and_preprocess(fis)

comms <- c('BCI')
comm_data <- fetch_and_preprocess(comms)

# SPDR  Gold Share ETF
gold <- c('GLD')
gold_data <- fetch_and_preprocess(gold)

# Long volatility
long_vol <- c('VIXY')
long_vol_data <- fetch_and_preprocess(long_vol)
```

### Portfolio 
```{r}
# Create a sample portfolio of different assets
portfolio <- cbind(stock_data, etf_data, fi_data, comm_data, gold_data)
portfolio_na_counts <- colSums(is.na(portfolio))
portfolio_na_counts
plot(portfolio, main = "Adjusted Closing Prices", xlab = "Date", ylab = "Price", col = rainbow(ncol(portfolio)), legend.loc = "topright")

# Convert the data from adjusted closing price to daily return
portfolio_dr <- Return.calculate(portfolio, method = "log")
portfolio_dr <- portfolio_dr * 100
# Remove first day w/o return
portfolio_dr <- portfolio_dr[-1, ]
# portfolio_dr_dis <- Return.calculate(portfolio, method = "discrete")
# Remove first day w/o return
# portfolio_dr_dis <- portfolio_dr_dis[-1, ]
assets <- colnames(portfolio_dr)
annual_rf = 0.04
```

### Common Constraints 
1. Full investment (weights sum to 1)  
2. Long-only = no short positions  

### Minimum Variance Portfolio
```{r}
# https://github.com/braverock/PortfolioAnalytics/blob/master/demo/demo_min_StdDev.R
port_spec_mv <- portfolio.spec(assets=assets)
port_spec_mv <- add.constraint(portfolio=port_spec_mv, type="full_investment")
port_spec_mv <- add.constraint(portfolio=port_spec_mv, type="long_only")
port_spec_mv <- add.objective(portfolio=port_spec_mv, type="risk", name="StdDev")
print(port_spec_mv)
min_var_portfolio <- optimize.portfolio(R=portfolio_dr, portfolio=port_spec_mv, 
                                       optimize_method="ROI", 
                                       trace=TRUE)
print(min_var_portfolio)
plot(min_var_portfolio, risk.col="StdDev", main="Long Only Minimize Portfolio StdDev")
```

### Markowitz Portfolio
```{r}
# https://github.com/braverock/PortfolioAnalytics/blob/master/demo/demo_max_Sharpe.R
port_spec_sr <- portfolio.spec(assets=assets)
port_spec_sr <- add.constraint(portfolio=port_spec_sr, type="full_investment")
port_spec_sr <- add.constraint(portfolio=port_spec_sr, type="long_only")
port_spec_sr <- add.objective(portfolio=port_spec_sr, type="return", name="mean")
port_spec_sr <- add.objective(portfolio=port_spec_sr, type="risk", name="StdDev")
max_sr_portfolio <- optimize.portfolio(R=portfolio_dr, portfolio=port_spec_sr, 
                                   optimize_method="ROI", 
                                   maxSR=TRUE, trace=TRUE, rf=annual_rf)
print(max_sr_portfolio)
plot(max_sr_portfolio, risk.col="StdDev", main="Long Only Max Sharpe Ratio StdDev")
```

### Comparison of Portfolios
```{r}
extractObjectiveMeasures(min_var_portfolio)
extractObjectiveMeasures(max_sr_portfolio)
extractWeights(min_var_portfolio)
extractWeights(max_sr_portfolio)
chart.Weights(min_var_portfolio)
chart.Weights(max_sr_portfolio)
extractStats(min_var_portfolio)
extractStats(max_sr_portfolio)
evaluate_portfolio <- function(portfolio_dr, weights) {
  portfolio_returns <- Return.portfolio(portfolio_dr, weights=weights)
  mean_return <- mean(portfolio_returns)
  std_dev <- sd(portfolio_returns)
  
  ann_mean_return <- mean_return * 252  # Assuming 252 trading days per year
  ann_std_dev <- std_dev * sqrt(252)    # Assuming 252 trading days per year
  
  # Calculate risk-adjusted measures
  sharpe_ratio <- ann_mean_return / ann_std_dev
  sortino_ratio <- SortinoRatio(portfolio_returns)
  
  evaluation_metrics <- data.frame(
    Mean_Return = mean_return,
    Std_Deviation = std_dev,
    Annualized_Mean_Return = ann_mean_return,
    Annualized_Std_Deviation = ann_std_dev,
    Sharpe_Ratio = sharpe_ratio,
    Sortino_Ratio = sortino_ratio
  )
  
  return(evaluation_metrics)
}

evaluate_portfolio(portfolio_dr, min_var_portfolio$weights)
evaluate_portfolio(portfolio_dr, max_sr_portfolio$weights)
```

### Consideration for Retirement Planning
#### Additional Constraints
1. Diversification: Diversification is important for retirement planning  
2. Conditional Value-at-Risk: Help control risk of large loss  
3. Specify target return of 6% which is slightly higher than current risk-free rate of 4%  

#### Risk Metric
1. Average Length (Retiree cannot rely on portfolio with long drawdown duration)  

#### Return Metric
1. Omega Sharpe Ratio  (Focus on downside risk)
2. Burke Ratio (Maximum Drowdown is important for retiree as they are dependent on withdrawals)  
3. Sortino Ratio (Picked to maximize capital preservation)  
```{r}
# Diversity - rebalancing over time? 
rp <- portfolio.spec(assets = assets)
rp <- add.constraint(rp, type="full_investment")
rp$constraints[[1]]$min_sum=0.99
rp$constraints[[1]]$max_sum=1.01
rp <- add.constraint(rp, type = "long_only")
rp <- add.constraint(rp, type = "CVaR", level = 0.95, enabled = TRUE)
rp <- add.constraint(rp, type = "return", return_target = 0.06, enabled = TRUE)
rp <- add.constraint(rp, type = "diversification", div_target = 0.75)

CumOmegaSharpeRatio <- function(R, weights, MAR=0) {
  portfolio_returns <- Return.portfolio(R, weights = weights)
  omega_sharpe_ratio <- OmegaSharpeRatio(portfolio_returns, MAR = MAR)
  return(omega_sharpe_ratio)
}

# Both Maximum Drawdown and Drawdown duration are important for retiree as they are dependent on withdrawals and further
# can't rely on portfolio with long drawdown duration
# Burke Ratio (return)
CumBurkeRatio <- function(R, weights) { 
  portfolio_returns <- Return.portfolio(R, weights = weights)
  burke_ratio <- BurkeRatio(portfolio_returns)
  return(burke_ratio)
}
# Average Length
CumAverageLength <- function(R, weights) {
  portfolio_returns <- Return.portfolio(R, weights = weights)
  average_length <- AverageLength(portfolio_returns)
  return(average_length)
}

rp <- add.objective(rp, type = "risk", name = 'CumAverageLength')
rp <- add.objective(rp, type = "return", name = 'CumOmegaSharpeRatio')
# rp <- add.objective(rp, type = 'return', name = 'CumBurkeRatio')

# Sortino Ratio is picked to maximize capital preservation (return)
rp <- add.objective(rp, type = "return", name = "SortinoRatio")
rp_portfolio <- optimize.portfolio(R=portfolio_dr, portfolio=rp, rf=annual_rf, search_size=1000)
rp_portfolio
```
### Comparisons 
```{r}
evaluate_portfolio(portfolio_dr, rp_portfolio$weights)
```

1. We can see that Minimum Variance Portfolio indeed produce lowest standard deviation but also lower mean return compared to Markowitz and our Retirement Portfolio  
2. The retirement portfolio provides higher mean return at the expense of higher risk in standard deviation.
3. The choice of portfolios however depend on the risk aversion and preferences of the manager  

### References
1. [https://bookdown.org/compfinezbook/introcompfinr/]
2. [https://rpubs.com/Sergio_Garcia/intermediate_portfolio_analysis_r]